---
title: "Cell line analysis"
author: "George Howitt"
date: "2023-05-22"
output: html_document
---

```{r}
#Load libraries
suppressPackageStartupMessages({
library(here)
library(BiocStyle)
library(dplyr)
library(janitor)
library(ggplot2)
library(cowplot)
library(patchwork)
library(DropletUtils)
library(tidyverse)
library(scuttle)
library(scater)
library(Seurat)
library(pheatmap)
library(speckle)
library(dittoSeq)
library(cellhashR)
library(RColorBrewer)
library(demuxmix)
library(ComplexHeatmap)
library(tidyHeatmap)
library(viridis)
}
)
```

# Cell line data.

This notebook contains the analysis code for all results and figures relating to the cell line data set in the paper "Benchmarking single-cell hashtag oligo demultiplexing methods".

The data set consists of one batch with three genetically distinct samples in each.
The batch was processed in three separate captures. We run all the demultiplexing methods on a per-capture level before then recombining for later analysis.

## Data loading and reduction.

Load in counts matrices and genetic IDs.
```{r}
lmo_counts_c1 <- read.csv(here("data", "cell_line_data", "lmo_counts_capture1.csv"), check.names = FALSE, row.names = 1)
lmo_counts_c2 <- read.csv(here("data", "cell_line_data", "lmo_counts_capture2.csv"), check.names = FALSE, row.names = 1)
lmo_counts_c3 <- read.csv(here("data", "cell_line_data", "lmo_counts_capture3.csv"), check.names = FALSE, row.names = 1)

lmo_donors_c1 <- read.csv(here("data", "cell_line_data", "lmo_donors_capture1.csv"), check.names = FALSE, row.names = 1)
lmo_donors_c2 <- read.csv(here("data", "cell_line_data", "lmo_donors_capture2.csv"), check.names = FALSE, row.names = 1)
lmo_donors_c3 <- read.csv(here("data", "cell_line_data", "lmo_donors_capture3.csv"), check.names = FALSE, row.names = 1)
```

Lists associating the HTOs with the genetic donors.
```{r}
LMO_list <- c("CL 01", "CL 02", "CL 03", "Doublet", "Negative")
donor_LMO_list <- list("CL A" = "CL 01", 
                       "CL B" = "CL 02", 
                       "CL C" = "CL 03", 
                       "Doublet" = "Doublet", 
                       "Negative" = "Negative")
LMO_donor_list <- list("CL 01" = "CL A", 
                       "CL 02" = "CL B", 
                       "CL 03" = "CL C", 
                       "Doublet" = "Doublet", 
                       "Negative" = "Negative")
```

Create Seurat objects
```{r}
seu_lmo_c1 <- CreateSeuratObject(counts = lmo_counts_c1, assay = "HTO")
seu_lmo_c2 <- CreateSeuratObject(counts = lmo_counts_c2, assay = "HTO")
seu_lmo_c3 <- CreateSeuratObject(counts = lmo_counts_c3, assay = "HTO")
```
```{r}
seu_lmo_c1$Barcode <- colnames(seu_lmo_c1)
seu_lmo_c2$Barcode <- colnames(seu_lmo_c2)
seu_lmo_c3$Barcode <- colnames(seu_lmo_c3)
```

Add genetic donor information to Seurat objects
```{r}
seu_lmo_c1$genetic_donor <- lmo_donors_c1$genetic_donor
seu_lmo_c2$genetic_donor <- lmo_donors_c2$genetic_donor
seu_lmo_c3$genetic_donor <- lmo_donors_c3$genetic_donor
```

Merge together for QC comparison
```{r}
seu_lmo <- merge(seu_lmo_c1, c(seu_lmo_c2, seu_lmo_c3))
```

Run PCAs and tSNEs
```{r}
DefaultAssay(seu_lmo_c1) <- "HTO"
seu_lmo_c1 <- NormalizeData(seu_lmo_c1, assay = "HTO", normalization.method = "CLR")
seu_lmo_c1 <- ScaleData(seu_lmo_c1, features = rownames(seu_lmo_c1),
    verbose = FALSE)
seu_lmo_c1 <- RunPCA(seu_lmo_c1, features = rownames(seu_lmo_c1), approx = FALSE, verbose = FALSE)
#seu_lmo_c1 <- RunTSNE(seu_lmo_c1, dims = 1:3, perplexity = 100, check_duplicates = FALSE, verbose = FALSE)

DefaultAssay(seu_lmo_c2) <- "HTO"
seu_lmo_c2 <- NormalizeData(seu_lmo_c2, assay = "HTO", normalization.method = "CLR")
seu_lmo_c2 <- ScaleData(seu_lmo_c2, features = rownames(seu_lmo_c2),
    verbose = FALSE)
seu_lmo_c2 <- RunPCA(seu_lmo_c2, features = rownames(seu_lmo_c2), approx = FALSE, verbose = FALSE)
#seu_lmo_c2 <- RunTSNE(seu_lmo_c2, dims = 1:3, perplexity = 100, check_duplicates = FALSE, verbose = FALSE)

DefaultAssay(seu_lmo_c3) <- "HTO"
seu_lmo_c3 <- NormalizeData(seu_lmo_c3, assay = "HTO", normalization.method = "CLR")
seu_lmo_c3 <- ScaleData(seu_lmo_c3, features = rownames(seu_lmo_c3),
    verbose = FALSE)
seu_lmo_c3 <- RunPCA(seu_lmo_c3, features = rownames(seu_lmo_c3), approx = FALSE, verbose = FALSE)
#seu_lmo_c3 <- RunTSNE(seu_lmo_c3, dims = 1:3, perplexity = 100, check_duplicates = FALSE, verbose = FALSE)
```

```{r}
DefaultAssay(seu_lmo) <- "HTO"
seu_lmo <- NormalizeData(seu_lmo, assay = "HTO", normalization.method = "CLR")
seu_lmo <- ScaleData(seu_lmo, features = rownames(seu_lmo),
    verbose = FALSE)
seu_lmo <- RunPCA(seu_lmo, features = rownames(seu_lmo), approx = FALSE, verbose = FALSE)
seu_lmo <- RunTSNE(seu_lmo, dims = 1:3, perplexity = 100, check_duplicates = FALSE, verbose = FALSE)
```

## QC plots

Density plots per barcode. In ideal conditions the density of the hashtag counts should appear bimodal, with a lower peak corresponding to the background and the higher peak corresponding to the signal.
```{r}
df <- as.data.frame(t(seu_lmo[["HTO"]]@counts))
colnames(df) <- LMO_donor_list[colnames(df)]
df %>%
  pivot_longer(cols = starts_with("CL")) %>%
  mutate(logged = log(value + 1)) %>%
  ggplot(aes(x = logged)) +
  xlab("log(counts)") +
  xlim(0.1,8) +
  geom_density(adjust = 2) +
  facet_wrap(~name, scales = "fixed", ncol = 3)  -> p1

p1
```

```{r}
p2 <- DimPlot(seu_lmo, group.by = "genetic_donor") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        axis.line.x = element_blank(), axis.line.y = element_blank(),
        plot.title = element_blank())

p2
```
```{r}
(p1 | p2) + plot_annotation(tag_levels = 'a') & 
         theme(plot.title = element_text(face = "plain", size = 10),
               plot.tag = element_text(face = 'plain'))
```

```{r}
#ggsave("LMO_QC.png",
#       plot = (p1 | p2) + plot_annotation(tag_levels = 'a') & 
#         theme(plot.title = element_text(face = "plain", size = 10),
#               plot.tag = element_text(face = 'plain')),
#       device = "png",
#       path = here("paper_latex", "figures"),
#       width = 10, height = 4,
#       units = "in",
#       dpi = 300)
```

## Hashtag-based demultiplexing

### hashedDrops
This function creates a list of hashedDrops calls. Its defaults are the same as hashedDrops
```{r}
create_hashedDrops_factor <- function(seurat_object, confident.min = 2,
                              doublet.nmads = 3, doublet.min = 2) {
  hto_counts <- GetAssayData(seurat_object[["HTO"]], slot = "counts")
  hash_stats <- DropletUtils::hashedDrops(hto_counts, confident.min = confident.min,
                                          doublet.nmads = doublet.nmads, doublet.min = doublet.min)
  
  hash_stats$Best <- rownames(seurat_object[["HTO"]])[hash_stats$Best]
  hash_stats$Second <- rownames(seurat_object[["HTO"]])[hash_stats$Second]
  
  HTO_assignments <- factor(case_when(
    hash_stats$Confident == TRUE ~ hash_stats$Best,
    hash_stats$Doublet == TRUE ~ "Doublet",
    TRUE ~ "Negative"))
  return(HTO_assignments)
  }
```

Making factors with "best" parameters
```{r}
seu_lmo_c1$hashedDrops_calls <- create_hashedDrops_factor(seu_lmo_c1, confident.min = 0.5)
seu_lmo_c2$hashedDrops_calls <- create_hashedDrops_factor(seu_lmo_c2, confident.min = 0.5)
seu_lmo_c3$hashedDrops_calls <- create_hashedDrops_factor(seu_lmo_c3, confident.min = 0.5)
```

Now with default parameters
```{r}
seu_lmo_c1$hashedDrops_default_calls <- create_hashedDrops_factor(seu_lmo_c1)
seu_lmo_c2$hashedDrops_default_calls <- create_hashedDrops_factor(seu_lmo_c2)
seu_lmo_c3$hashedDrops_default_calls <- create_hashedDrops_factor(seu_lmo_c3)
```

### HashSolo

HashSolo is a scanpy program. Needs a bit of prep
Write to anndata compatible files
Counts
```{r}
library(Matrix)
writeMM(seu_lmo_c1@assays$HTO@counts, here("data", "cell_line_data", "adata", "c1_counts.mtx"))
writeMM(seu_lmo_c2@assays$HTO@counts, here("data", "cell_line_data", "adata", "c2_counts.mtx"))
writeMM(seu_lmo_c3@assays$HTO@counts, here("data", "cell_line_data", "adata", "c3_counts.mtx"))
```
Barcodes
```{r}
barcodes <- data.frame(colnames(seu_lmo_c1))
colnames(barcodes)<-'Barcode'
write.csv(barcodes, here("data", "cell_line_data", "adata", "c1_barcodes.csv"),
          quote = FALSE,row.names = FALSE)
barcodes <- data.frame(colnames(seu_lmo_c2))
colnames(barcodes)<-'Barcode'
write.csv(barcodes, here("data", "cell_line_data", "adata", "c2_barcodes.csv"),
          quote = FALSE,row.names = FALSE)
barcodes <- data.frame(colnames(seu_lmo_c3))
colnames(barcodes)<-'Barcode'
write.csv(barcodes, here("data", "cell_line_data", "adata", "c3_barcodes.csv"),
          quote = FALSE,row.names = FALSE)
```
Save LMO names (just need one per capture)
```{r}
HTOs <- data.frame(rownames(seu_lmo_c1))
colnames(HTOs) <- 'HTO'
write.csv(HTOs, here("data", "cell_line_data", "adata", "HTOs.csv"),
          quote = FALSE,row.names = FALSE)
```

See hashsolo_calls.ipynb for how we get these assignments
```{r}
seu_lmo_c1$hashsolo_calls <- read.csv(here("data", "cell_line_data", "adata", "c1_hashsolo.csv"))$Classification
seu_lmo_c2$hashsolo_calls <- read.csv(here("data", "cell_line_data", "adata", "c2_hashsolo.csv"))$Classification
seu_lmo_c3$hashsolo_calls <- read.csv(here("data", "cell_line_data", "adata", "c3_hashsolo.csv"))$Classification
```

### HTODemux
```{r}
HDmux <- HTODemux(seu_lmo_c1)
seu_lmo_c1$HTODemux_calls <- HDmux$hash.ID
HDmux <- HTODemux(seu_lmo_c2)
seu_lmo_c2$HTODemux_calls <- HDmux$hash.ID
HDmux <- HTODemux(seu_lmo_c3)
seu_lmo_c3$HTODemux_calls <- HDmux$hash.ID
```

###GMM-Demux

GMM-Demux is run on the command line and needs a function to read in the results and format them all properly.
```{r}
create_gmm_demux_factor <- function(seu, GMM_path, hto_list) {
  #Read in output, have to use the "full" report, not the simplified one.
  calls <- read.csv(paste0(GMM_path, "/GMM_full.csv"), row.names = 1)
  #Read in names of clusters
  cluster_names <- read.table(paste0(GMM_path, "/GMM_full.config"), sep = ",")
  names(cluster_names) <- c("Cluster_id", "assignment")
  #Need to fix the formatting of the assignment names, for some reason there's a leading space.
  cluster_names$assignment <- gsub(x = cluster_names$assignment, pattern = '^ ', replacement = '')
  #Add cell barcodes 
  calls$Barcode <- rownames(calls)
  calls <- merge(calls, cluster_names, by = "Cluster_id", sort = FALSE)
  #Need to re-order after merge for some reason
  calls <- calls[order(match(calls$Barcode, names(seu$Barcode))), ]
  #Rename the negative cluster for consistency
  calls$assignment[calls$assignment == "negative"] <- "Negative"
  #Put all the multiplet states into one assignment category
  calls$assignment[!calls$assignment %in% c("Negative", hto_list)] <- "Doublet"
  return(as.factor(calls$assignment))
}
```

Need to write transpose of counts matrices to .csv files to run GMM-Demux on command line. 
```{r}
write.csv(t(as.matrix(lmo_counts_c1)), here("data", "cell_line_data", "GMM-Demux", "c1_hto_counts_transpose.csv"))
write.csv(t(as.matrix(lmo_counts_c2)), here("data", "cell_line_data", "GMM-Demux", "c2_hto_counts_transpose.csv"))
write.csv(t(as.matrix(lmo_counts_c3)), here("data", "cell_line_data", "GMM-Demux", "c3_hto_counts_transpose.csv"))
```

See script for running GMM-Demux

Add to objects
```{r}
seu_lmo_c1$GMMDemux_calls <- create_gmm_demux_factor(seu_lmo_c1, here("data", "cell_line_data", "GMM-Demux", "gmm_out_cell_line_c1", "full_report"), LMO_list) 
seu_lmo_c2$GMMDemux_calls <- create_gmm_demux_factor(seu_lmo_c2, here("data", "cell_line_data", "GMM-Demux", "gmm_out_cell_line_c2", "full_report"), LMO_list) 
seu_lmo_c3$GMMDemux_calls <- create_gmm_demux_factor(seu_lmo_c3, here("data", "cell_line_data", "GMM-Demux", "gmm_out_cell_line_c3", "full_report"), LMO_list)
```

###deMULTIplex

Next is deMULTIplex, using the Seurat wrapper function MULTIseqDemux for this
```{r}
seu_lmo_c1$deMULTIplex_calls <- MULTIseqDemux(seu_lmo_c1, autoThresh = TRUE)$MULTI_ID
seu_lmo_c2$deMULTIplex_calls <- MULTIseqDemux(seu_lmo_c2, autoThresh = TRUE)$MULTI_ID
seu_lmo_c3$deMULTIplex_calls <- MULTIseqDemux(seu_lmo_c3, autoThresh = TRUE)$MULTI_ID
```

###BFF
Finally cellhashR's BFF_raw and BFF_cluster methods.
Need to run this on the raw counts matrix
```{r}
lmo_counts_c1 <- seu_lmo_c1[["HTO"]]@counts
lmo_counts_c2 <- seu_lmo_c2[["HTO"]]@counts
lmo_counts_c3 <- seu_lmo_c3[["HTO"]]@counts
```

```{r}
cellhashR_calls <- GenerateCellHashingCalls(barcodeMatrix = lmo_counts_c1, methods = c("bff_raw", "bff_cluster"), doTSNE = FALSE, doHeatmap = FALSE)
seu_lmo_c1$BFF_raw_calls <- cellhashR_calls$bff_raw
seu_lmo_c1$BFF_cluster_calls <- cellhashR_calls$bff_cluster

cellhashR_calls <- GenerateCellHashingCalls(barcodeMatrix = lmo_counts_c2, methods = c("bff_raw", "bff_cluster"), doTSNE = FALSE, doHeatmap = FALSE)
seu_lmo_c2$BFF_raw_calls <- cellhashR_calls$bff_raw
seu_lmo_c2$BFF_cluster_calls <- cellhashR_calls$bff_cluster

cellhashR_calls <- GenerateCellHashingCalls(barcodeMatrix = lmo_counts_c3, methods = c("bff_raw", "bff_cluster"), doTSNE = FALSE, doHeatmap = FALSE)
seu_lmo_c3$BFF_raw_calls <- cellhashR_calls$bff_raw
seu_lmo_c3$BFF_cluster_calls <- cellhashR_calls$bff_cluster
```

###demuxmix

This function turns the output of demuxmix into something consistent with the other methods
```{r}
demuxmix_calls_consistent <- function(seurat_object, model = "naive", hto_list) {
  hto_counts <- as.matrix(GetAssayData(seurat_object[["HTO"]], slot = "counts"))
  dmm <- demuxmix(hto_counts, model = model)
  dmm_calls <- dmmClassify(dmm)
  calls_out <- case_when(dmm_calls$HTO %in% hto_list ~ dmm_calls$HTO,
               !dmm_calls$HTO %in% hto_list ~ case_when(
                 dmm_calls$Type == "multiplet" ~ "Doublet",
                 dmm_calls$Type %in% c("negative", "uncertain") ~ "Negative")
               )
  return(as.factor(calls_out))
}
```

```{r}
seu_lmo_c1$demuxmix_calls <- demuxmix_calls_consistent(seu_lmo_c1, hto_list = LMO_list)
seu_lmo_c2$demuxmix_calls <- demuxmix_calls_consistent(seu_lmo_c2, hto_list = LMO_list)
seu_lmo_c3$demuxmix_calls <- demuxmix_calls_consistent(seu_lmo_c3, hto_list = LMO_list)
```

Re-merge back into single seurat object
```{r}
seu_lmo_c1$capture <- "capture 1"
seu_lmo_c2$capture <- "capture 2"
seu_lmo_c3$capture <- "capture 3"

seu_lmo <- merge(seu_lmo_c1, c(seu_lmo_c2, seu_lmo_c3))
```

Save Seurat objects with all the hashtag assignments
```{r}
saveRDS(seu_lmo, here("data", "cell_line_data", "lmo_all_methods.SEU.rds"))
```

## Making plots

### Compute F-scores

We compute the F-score of each of the possible singlet assignments. 
```{r}
#Helper function
calculate_HTO_fscore <- function(seurat_object, donor_hto_list, method) {
  calls <- seurat_object[[method]]
  f <- NULL
  for (HTO in donor_hto_list) {
    tp <- sum(calls == HTO & donor_hto_list[seurat_object$genetic_donor] == HTO) #True positive rate
    fp <- sum(calls == HTO & donor_hto_list[seurat_object$genetic_donor] != HTO) #False positive rate
    fn <- sum(calls != HTO & donor_hto_list[seurat_object$genetic_donor] == HTO) #False negative rate
    f <- c(f, tp / (tp + 0.5 * (fp + fn)))
  }
#  f <- c(f, median(f)) #Add median F score
  f <- c(f, mean(f)) #Add mean F score

  names(f) <- c(donor_hto_list, "Average")
  
  return(f)
}
```

Compare F scores for the methods.
```{r}
Fscore_hashedDrops <- calculate_HTO_fscore(seu_lmo, donor_LMO_list[1:3], "hashedDrops_calls")
Fscore_hashedDrops_default <- calculate_HTO_fscore(seu_lmo, donor_LMO_list[1:3], "hashedDrops_default_calls")
Fscore_HTODemux <- calculate_HTO_fscore(seu_lmo, donor_LMO_list[1:3], "HTODemux_calls")
Fscore_GMMDemux <- calculate_HTO_fscore(seu_lmo, donor_LMO_list[1:3], "GMMDemux_calls")
Fscore_deMULTIplex <- calculate_HTO_fscore(seu_lmo, donor_LMO_list[1:3], "deMULTIplex_calls")
Fscore_BFF_raw <- calculate_HTO_fscore(seu_lmo, donor_LMO_list[1:3], "BFF_raw_calls")
Fscore_BFF_cluster <- calculate_HTO_fscore(seu_lmo, donor_LMO_list[1:3], "BFF_cluster_calls")
Fscore_demuxmix <- calculate_HTO_fscore(seu_lmo, donor_LMO_list[1:3], "demuxmix_calls")
Fscore_hashsolo <- calculate_HTO_fscore(seu_lmo, donor_LMO_list[1:3], "hashsolo_calls")
```

```{r}
Fscore_matrix <- data.frame("LMO" = c(LMO_list[1:3], "Mean"),
                            "hashedDrops" = Fscore_hashedDrops,
                            "hashedDrops_default" = Fscore_hashedDrops_default,
                            "HashSolo" = Fscore_hashsolo,
                            "HTODemux" = Fscore_HTODemux,
                            "GMM_Demux" = Fscore_GMMDemux,
                            "deMULTIplex" = Fscore_deMULTIplex,
                            "BFF_raw" = Fscore_BFF_raw,
                            "BFF_cluster" = Fscore_BFF_cluster,
                            "demuxmix" = Fscore_demuxmix)

#Removing average information for this data set
Fscore_matrix = Fscore_matrix[1:3,]
```

```{r}
Fscore_matrix %>% 
  pivot_longer(cols = c("hashedDrops", "hashedDrops_default", "HashSolo", "HTODemux", "GMM_Demux", "deMULTIplex", "BFF_raw", "BFF_cluster", "demuxmix"),
               names_to = "method",
               values_to = "Fscore") -> Fscore_matrix
```

```{r}
p1 <- heatmap(Fscore_matrix, 
        .row = method, 
        .column = LMO, 
        .value = Fscore,
        column_title = "F-score - Cell line data",
        cluster_rows = TRUE,
        row_names_gp = gpar(fontsize = 10),
        show_row_dend = FALSE,
        row_names_side = "left", 
        row_title = "",
        cluster_columns = FALSE,
        column_names_gp = gpar(fontsize = 10),
        palette_value = plasma(3)) %>%
  wrap_heatmap()

p1
```
```{r}
#ggsave(here("paper_latex", "figures", "CL_Fscore.png"),
#       p1,
#       device = "png",
#      width = 6, height = 4,
#       units = "in",
#       dpi = 350
#      )
```

### Doublet-to-singlet assignments

```{r}
method_calls <- c("hashedDrops_calls", 
                  "hashedDrops_default_calls", 
                  "hashsolo_calls", 
                  "HTODemux_calls", 
                  "GMMDemux_calls",
                  "deMULTIplex_calls", 
                  "BFF_raw_calls", 
                  "BFF_cluster_calls", 
                  "demuxmix_calls")

doublets <- seu_lmo[, seu_lmo$genetic_donor == "Doublet"]
doublet_doublet <- NULL
doublet_negative <- NULL
doublet_singlet <- NULL

for (method in method_calls) {
  doublet_doublet <- c(doublet_doublet, sum(seu_lmo$genetic_donor == "Doublet" & seu_lmo[[method]] == "Doublet") / sum(seu_lmo$genetic_donor == "Doublet"))
  doublet_negative <- c(doublet_negative, sum(seu_lmo$genetic_donor == "Doublet" & seu_lmo[[method]] == "Negative") / sum(seu_lmo$genetic_donor == "Doublet"))
  doublet_singlet <- c(doublet_singlet, sum(seu_lmo$genetic_donor == "Doublet" & Reduce("|", lapply(LMO_list[1:3], function(x) seu_lmo[[method]] == x))) / sum(seu_lmo$genetic_donor == "Doublet"))
}

names(doublet_doublet) <- method_calls
names(doublet_negative) <- method_calls
names(doublet_singlet) <- method_calls
```

```{r}
doublet_assignments <- data.frame("method" = c("hashedDrops", 
                                               "hashedDrops (default)", 
                                               "HashSolo", "HTODemux",
                                               "GMM-Demux", 
                                               "deMULTIplex", 
                                               "BFF_raw",
                                               "BFF_cluster", 
                                               "demuxmix"),
                            "Doublets" = doublet_doublet,
                            "Negative" = doublet_negative,
                            "Singlet" = doublet_singlet) %>%
  pivot_longer(cols = c("Doublets", "Negative", "Singlet"), 
               names_to = "assignment",
               values_to = "fraction")
```


```{r}
doublet_colours <- c("black", "gray60", "firebrick1")

p2 <- ggplot(doublet_assignments %>%
               mutate(method = factor(method, levels = c("hashedDrops (default)", 
                                                         "HTODemux", 
                                                         "demuxmix", 
                                                         "BFF_raw", 
                                                         "HashSolo", 
                                                         "GMM-Demux", 
                                                         "deMULTIplex", 
                                                         "hashedDrops", 
                                                         "BFF_cluster")))) + 
  geom_bar(aes(x = method, y = fraction, fill = assignment),
           stat = "identity") +
  ggtitle("Cell line data (4945 doublets)") +
  ylim(0, 1) +
  scale_fill_manual(values = doublet_colours) +
  theme(axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(size = 10)
        ) + coord_flip()

p2
```

```{r}
#ggsave(here("paper_latex", "figures", "CL_doublet_assignments.png"),
#       p2,
#       device = "png",
#       width = 6, height = 4,
#       units = "in",
#       dpi = 350
#      )
```

How many droplets in the third peak?
```{r}
#df <- as.data.frame(t(seu_lmo[["HTO"]]@counts))
#colnames(df) <- gsub("_", " ", LMO_donor_list[colnames(df)])
#df %>%
#  pivot_longer(cols = starts_with("donor")) %>%
#  mutate(logged = log10(value + 1)) %>%
#  ggplot(aes(x = logged)) +
# xlab("log10(counts)") +
#  xlim(0.1,4) +
#  geom_density(adjust = 2) +
#  facet_wrap(~name, scales = "fixed", ncol = 3)  -> p1

#p1
```
```{r}
#table(seu_lmo$genetic_donor)
#sum(log10(seu_lmo[["HTO"]]@counts[1,]) > 2)
#sum(log10(seu_lmo[["HTO"]]@counts[2,]) > 2.5)
#sum(log10(seu_lmo[["HTO"]]@counts[3,]) > 2.5)
```













