---
title: "Ovarian tissue data"
author: "George Howitt"
date: "2023-05-16"
output: html_document
---

```{r}
#Load libraries
suppressPackageStartupMessages({
library(here)
library(BiocStyle)
library(dplyr)
library(janitor)
library(ggplot2)
library(cowplot)
library(patchwork)
library(DropletUtils)
library(tidyverse)
library(scuttle)
library(scater)
library(Seurat)
library(pheatmap)
library(speckle)
library(dittoSeq)
library(cellhashR)
library(RColorBrewer)
library(demuxmix)
library(ComplexHeatmap)
library(tidyHeatmap)
library(viridis)
})
```

# Ovarian tissue data
This notebook contains the analysis code for all results and figures relating to the cell line data set in the paper "Benchmarking single-cell hashtag oligo demultiplexing methods".

This data consists of eight samples where scRNA-seq was performed in two batches of four samples each. on ovarian tumours. 

## Data loading and reduction

Data is available here https://www.ncbi.nlm.nih.gov/projects/gap/cgi-bin/study.cgi?study_id=phs002262.v2.p1&phv=436020&phd=&pha=&pht=10575&phvf=&phdf=&phaf=&phtf=&dssp=1&consent=&temp=1

```{r}
hashtag_counts_b1 <- read.csv(here("data", "solid_tumor_data", "ovarian_tumour_b1.csv"),
                                check.names = FALSE, row.names = 1)
hashtag_counts_b2 <- read.csv(here("data", "solid_tumor_data", "ovarian_tumour_b2.csv"), 
                                check.names = FALSE, row.names = 1)
donors_b1 <- read.csv(here("data", "solid_tumor_data", "ovarian_tumour_donors_b1.csv"), 
                      row.names = 1)
donors_b2 <- read.csv(here("data", "solid_tumor_data", "ovarian_tumour_donors_b2.csv"), 
                      row.names = 1)
```


```{r}
seu_b1 <- CreateSeuratObject(counts = hashtag_counts_b1, assay = "HTO")
seu_b2 <- CreateSeuratObject(counts = hashtag_counts_b2, assay = "HTO")
```

```{r}
seu_b1$Barcode <- colnames(seu_b1)
seu_b2$Barcode <- colnames(seu_b2)
```

Add genetic donor information to Seurat objects
```{r}
seu_b1$genetic_donor <- donors_b1$genetic_donors
seu_b2$genetic_donor <- donors_b2$genetic_donors
```

Useful lists
```{r}
hashtag_list_b1 <- c("OT 01",
                     "OT 02",
                     "OT 03",
                     "OT 04")

donor_hashtag_list_b1 <- list("OT A" = "OT 01", 
                              "OT B" = "OT 02",
                              "OT C" = "OT 03",
                              "OT D" = "OT 04",
                              "Doublet" = "Doublet", 
                              "Negative" = "Negative")

hashtag_donor_list_b1 <- list("OT 01" = "OT A",
                              "OT 02" = "OT B",
                              "OT 03" = "OT C",
                              "OT 04" = "OT D",
                              "Doublet" = "Doublet", 
                              "Negative" = "Negative")

hashtag_list_b2 <- c("OT 05",
                     "OT 06",
                     "OT 07",
                     "OT 08")

donor_hashtag_list_b2 <- list("OT E" = "OT 05", 
                              "OT F" = "OT 06",
                              "OT G" = "OT 07",
                              "OT H" = "OT 08",
                              "Doublet" = "Doublet", 
                              "Negative" = "Negative")

hashtag_donor_list_b2 <- list("OT 05" = "OT E",
                              "OT 06" = "OT F",
                              "OT 07" = "OT G",
                              "OT 08" = "OT H",
                              "Doublet" = "Doublet", 
                              "Negative" = "Negative")
```

```{r}
DefaultAssay(seu_b1) <- "HTO"
seu_b1 <- NormalizeData(seu_b1, assay = "HTO", normalization.method = "CLR")
seu_b1 <- ScaleData(seu_b1, features = rownames(seu_b1),
    verbose = FALSE)
seu_b1 <- RunPCA(seu_b1, features = rownames(seu_b1), approx = FALSE, verbose = FALSE)
seu_b1 <- RunTSNE(seu_b1, dims = 1:3, perplexity = 100, check_duplicates = FALSE, verbose = FALSE)

DefaultAssay(seu_b2) <- "HTO"
seu_b2 <- NormalizeData(seu_b2, assay = "HTO", normalization.method = "CLR")
seu_b2 <- ScaleData(seu_b2, features = rownames(seu_b2),
    verbose = FALSE)
seu_b2 <- RunPCA(seu_b2, features = rownames(seu_b2), approx = FALSE, verbose = FALSE)
seu_b2 <- RunTSNE(seu_b2, dims = 1:3, perplexity = 100, check_duplicates = FALSE, verbose = FALSE)
```

## QC Plots

Density plots per barcode. In ideal conditions the density of the hashtag counts should appear bimodal, with a lower peak corresponding to the background and the higher peak corresponding to the signal.

```{r}
df <- as.data.frame(t(seu_b1[["HTO"]]@counts))
df %>%
  pivot_longer(cols = starts_with("OT")) %>%
  mutate(logged = log(value + 1)) %>%
  ggplot(aes(x = logged)) +
  xlab("log(counts)") +
  xlim(3,10) +
  geom_density(adjust = 2) +
  facet_wrap(~name, scales = "fixed", ncol = 2)  -> p1

df <- as.data.frame(t(seu_b2[["HTO"]]@counts))
df %>%
  pivot_longer(cols = starts_with("OT")) %>%
  mutate(logged = log(value + 1)) %>%
  ggplot(aes(x = logged)) +
  xlab("log(counts)") +
  xlim(3,10) +
  geom_density(adjust = 2) +
  facet_wrap(~name, scales = "fixed", ncol = 2)  -> p2

p1 / p2
```
These data aren't looking so hot. Hard to see a second peak in most hashtags...

```{r}
p3 <- DimPlot(seu_b1, group.by = "genetic_donor") + 
  ggtitle("Batch 1")

p4 <- DimPlot(seu_b2, group.by = "genetic_donor") + 
  ggtitle("Batch 2")

p3 / p4
```

```{r}
(((p1 / p2 ) | (p3 / p4)) + plot_annotation(tag_levels = 'a')) & 
  theme(plot.title = element_text(face = "plain", size = 10), 
        plot.tag = element_text(face = 'plain'))
```


# Demultiplexing

## hashedDrops
This function creates a list of hashedDrops calls. Its defaults are the same as hashedDrops
```{r}
create_hashedDrops_factor <- function(seurat_object, confident.min = 2,
                              doublet.nmads = 3, doublet.min = 2) {
  hto_counts <- GetAssayData(seurat_object[["HTO"]], slot = "counts")
  hash_stats <- DropletUtils::hashedDrops(hto_counts, confident.min = confident.min,
                                          doublet.nmads = doublet.nmads, doublet.min = doublet.min)
  
  hash_stats$Best <- rownames(seurat_object[["HTO"]])[hash_stats$Best]
  hash_stats$Second <- rownames(seurat_object[["HTO"]])[hash_stats$Second]
  
  HTO_assignments <- factor(case_when(
    hash_stats$Confident == TRUE ~ hash_stats$Best,
    hash_stats$Doublet == TRUE ~ "Doublet",
    TRUE ~ "Negative"))
  return(HTO_assignments)
  }
```

Making factors with best parameters
```{r}
seu_b1$hashedDrops_calls <- create_hashedDrops_factor(seu_b1, confident.min = 0.5)
seu_b2$hashedDrops_calls <- create_hashedDrops_factor(seu_b2, confident.min = 0.5)
```

Now with default parameters
```{r}
seu_b1$hashedDrops_default_calls <- create_hashedDrops_factor(seu_b1)
seu_b2$hashedDrops_default_calls <- create_hashedDrops_factor(seu_b2)
```

## HashSolo

HashSolo is a scanpy program. Needs a bit of prep
Write to anndata compatible files
Counts
```{r}
library(Matrix)
writeMM(seu_b1@assays$HTO@counts, here("data", "solid_tumor_data", "adata", "b1_counts.mtx"))
writeMM(seu_b2@assays$HTO@counts, here("data", "solid_tumor_data", "adata", "b2_counts.mtx"))
```
Barcodes
```{r}
barcodes <- data.frame(colnames(seu_b1))
colnames(barcodes)<-'Barcode'
write.csv(barcodes, here("data", "solid_tumor_data", "adata", "b1_barcodes.csv"),
          quote = FALSE,row.names = FALSE)
barcodes <- data.frame(colnames(seu_b2))
colnames(barcodes)<-'Barcode'
write.csv(barcodes, here("data", "solid_tumor_data", "adata", "b2_barcodes.csv"),
          quote = FALSE,row.names = FALSE)
```
Save hashtag names 
```{r}
HTOs_b1 <- data.frame(rownames(seu_b1))
colnames(HTOs_b1) <- 'HTO'
write.csv(HTOs_b1, here("data", "solid_tumor_data", "adata", "HTOs_b1.csv"),
          quote = FALSE,row.names = FALSE)
HTOs_b2 <- data.frame(rownames(seu_b2))
colnames(HTOs_b2) <- 'HTO'
write.csv(HTOs_b2, here("data", "solid_tumor_data", "adata", "HTOs_b2.csv"),
          quote = FALSE,row.names = FALSE)
```

See hashsolo_calls.ipynb for how we get these assignments
```{r}
seu_b1$hashsolo_calls <- read.csv(here("data", "solid_tumor_data", "adata",  "b1_hashsolo.csv"))$Classification
seu_b2$hashsolo_calls <- read.csv(here("data", "solid_tumor_data", "adata",  "b2_hashsolo.csv"))$Classification
```

## deMULTIplex
```{r}
seu_b1$deMULTIplex_calls <- MULTIseqDemux(seu_b1, autoThresh = TRUE)$MULTI_ID
seu_b2$deMULTIplex_calls <- MULTIseqDemux(seu_b2, autoThresh = TRUE)$MULTI_ID
```

## HTODemux
```{r}
HDmux <- HTODemux(seu_b1)
seu_b1$HTODemux_calls <- HDmux$hash.ID
HDmux <- HTODemux(seu_b2)
seu_b2$HTODemux_calls <- HDmux$hash.ID
```

## GMM-Demux

GMM-Demux is run on the command line and needs a function to read in the results and format them all properly.
```{r}
create_gmm_demux_factor <- function(seu, GMM_path, hto_list) {
  #Read in output, have to use the "full" report, not the simplified one.
  calls <- read.csv(paste0(GMM_path, "/GMM_full.csv"), row.names = 1)
  #Read in names of clusters
  cluster_names <- read.table(paste0(GMM_path, "/GMM_full.config"), sep = ",")
  names(cluster_names) <- c("Cluster_id", "assignment")
  #Need to fix the formatting of the assignment names, for some reason there's a leading space.
  cluster_names$assignment <- gsub(x = cluster_names$assignment, pattern = '^ ', replacement = '')
  #Add cell barcodes 
  calls$Barcode <- rownames(calls)
  calls <- merge(calls, cluster_names, by = "Cluster_id", sort = FALSE)
  #Need to re-order after merge for some reason
  calls <- calls[order(match(calls$Barcode, names(seu$Barcode))), ]
  #Rename the negative cluster for consistency
  calls$assignment[calls$assignment == "negative"] <- "Negative"
  #Put all the multiplet states into one assignment category
  calls$assignment[!calls$assignment %in% c("Negative", hto_list)] <- "Doublet"
  return(as.factor(calls$assignment))
}
```

Write the transpose of the counts matrix for GMM-Demux
```{r}
write.csv(t(hashtag_counts_b1), here("data", "solid_tumor_data", "GMM-Demux", "hashtag_counts_b1_transpose.csv"))
write.csv(t(hashtag_counts_b2), here("data", "solid_tumor_data", "GMM-Demux", "hashtag_counts_b2_transpose.csv"))
```

Run the script run_GMM_demux_solid_tumor.sh in command line.

Add to objects
```{r}
seu_b1$GMMDemux_calls <- create_gmm_demux_factor(seu_b1, here("data", "solid_tumor_data", "GMM-Demux", "gmm_out_b1", "full_report"), hashtag_list_b1) 
seu_b2$GMMDemux_calls <- create_gmm_demux_factor(seu_b2, here("data", "solid_tumor_data", "GMM-Demux", "gmm_out_b2", "full_report"), hashtag_list_b2) 
```

## BFF
```{r}
cellhashR_calls <- GenerateCellHashingCalls(barcodeMatrix = hashtag_counts_b1, 
                                            methods = c("bff_raw", "bff_cluster"), 
                                            doTSNE = FALSE, doHeatmap = FALSE)
seu_b1$BFF_raw_calls <- cellhashR_calls$bff_raw
seu_b1$BFF_cluster_calls <- cellhashR_calls$bff_cluster

cellhashR_calls <- GenerateCellHashingCalls(barcodeMatrix = hashtag_counts_b2, 
                                            methods = c("bff_raw", "bff_cluster"), 
                                            doTSNE = FALSE, doHeatmap = FALSE)
seu_b2$BFF_raw_calls <- cellhashR_calls$bff_raw
seu_b2$BFF_cluster_calls <- cellhashR_calls$bff_cluster
```

```{r}
seu_b1$BFF_cluster_calls <- case_when(seu_b1$BFF_cluster_calls == "Not Called" ~ "Negative")
```


## demuxmix

This function turns the output of demuxmix into something consistent with the other methods
```{r}
demuxmix_calls_consistent <- function(seurat_object, model = "naive", hto_list) {
  hto_counts <- as.matrix(GetAssayData(seurat_object[["HTO"]], slot = "counts"))
  dmm <- demuxmix(hto_counts, model = model)
  dmm_calls <- dmmClassify(dmm)
  calls_out <- case_when(dmm_calls$HTO %in% hto_list ~ dmm_calls$HTO,
               !dmm_calls$HTO %in% hto_list ~ case_when(
                 dmm_calls$Type == "multiplet" ~ "Doublet",
                 dmm_calls$Type %in% c("negative", "uncertain") ~ "Negative")
               )
  return(as.factor(calls_out))
}
```

```{r}
seu_b1$demuxmix_calls <- demuxmix_calls_consistent(seu_b1, hto_list = hashtag_list_b1)
seu_b2$demuxmix_calls <- demuxmix_calls_consistent(seu_b2, hto_list = hashtag_list_b2)
```

Save Seurat objects with all the hashtag assignments
```{r}
#saveRDS(seu_b1, here("data", "solid_tumor_data", "batch1_all_methods.SEU.rds"))
#saveRDS(seu_b2, here("data", "solid_tumor_data", "batch2_all_methods.SEU.rds"))
```

# Making plots

## Category plots

Looking at what fraction of droplets are assigned as singlets, doublets or negative by each method
```{r}
method_calls <- c("hashedDrops_calls", 
                  "hashedDrops_default_calls", 
                  "hashsolo_calls", 
                  "HTODemux_calls", 
                  "GMMDemux_calls",
                  "deMULTIplex_calls", 
                  "BFF_raw_calls", 
                  "BFF_cluster_calls", 
                  "demuxmix_calls")

for (method in method_calls){ 
  seu_b1[[gsub(method, paste0(method, "_donors"), method)]] <- as.factor(unlist(hashtag_list_b1[unlist(seu_b1[[method]])]))
  seu_b2[[gsub(method, paste0(method, "_donors"), method)]] <- as.factor(unlist(hashtag_list_b2[unlist(seu_b2[[method]])]))
}

seu_b1$Batch <- "Batch 1"
seu_b2$Batch <- "Batch 2"
seu <- merge(seu_b1, seu_b2)
seu$Batch <- as.factor(seu$Batch)
```
Need to make a factor which reduces the assignments to one of ("Singlet", "Doublet", "Negative"). 
```{r}
sd_or_u_hashtags <- function(seurat_object, method) {
  return(case_when(seurat_object[[method]] == "Doublet" ~ "Doublet",
            seurat_object[[method]] == "Negative" ~ "Negative",
            TRUE ~ "Singlet"))
}

for (method in method_calls) {
  seu[[gsub(method, paste0(method, "_category"), method)]] <- sd_or_u_hashtags(seu, method)
}
seu$genetic_donor_category <- case_when(seu$genetic_donor == "Doublet" ~ "Doublet",
                                        seu$genetic_donor == "Negative" ~ "Negative",
                                        TRUE ~ "Singlet")
```

```{r}
p1 <- dittoBarPlot(seu, var = "genetic_donor_category", group.by = "Batch") + NoLegend() +
  ggtitle("vireo (genetics)") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank())
p2 <- dittoBarPlot(seu, var = "BFF_cluster_calls_category", group.by = "Batch") + NoLegend() + 
  ggtitle("BFF_cluster") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.text.y = element_blank())
p3 <- dittoBarPlot(seu, var = "BFF_raw_calls_category", group.by = "Batch") + NoLegend() + 
  ggtitle("BFF_raw") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank())
p4 <- dittoBarPlot(seu, var = "deMULTIplex_calls_category", group.by = "Batch") + 
  ggtitle("deMULTIplex") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.text.y = element_blank())
p5 <- dittoBarPlot(seu, var = "demuxmix_calls_category", group.by = "Batch") + NoLegend() +
  ggtitle("demuxmix") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank())
p6 <- dittoBarPlot(seu, var = "GMMDemux_calls_category", group.by = "Batch") + NoLegend() + 
  ggtitle("GMM-Demux") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.text.y = element_blank())
p7 <- dittoBarPlot(seu, var = "hashedDrops_calls_category", group.by = "Batch") + NoLegend() + 
  ggtitle("hashedDrops - best") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank())
p8 <- dittoBarPlot(seu, var = "hashedDrops_default_calls_category", group.by = "Batch") + NoLegend() + 
  ggtitle("hashedDrops - default") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.text.y = element_blank())
p9 <- dittoBarPlot(seu, var = "hashsolo_calls_category", group.by = "Batch") + NoLegend() +
  ggtitle("HashSolo") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14))
p10 <- dittoBarPlot(seu, var = "HTODemux_calls_category", group.by = "Batch") + NoLegend() +
  ggtitle("HTODemux") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14), axis.text.y = element_blank())

(p1 + p2) / (p3 + p4) / (p5 + p6) / (p7 + p8) / (p9 + p10)
```

## Fscore

```{r}
calculate_HTO_fscore <- function(seurat_object, donor_hto_list, method) {
  calls <- seurat_object[[method]]
  f <- NULL
  for (HTO in donor_hto_list) {
    tp <- sum(calls == HTO & donor_hto_list[seurat_object$genetic_donor] == HTO) #True positive rate
    fp <- sum(calls == HTO & donor_hto_list[seurat_object$genetic_donor] != HTO) #False positive rate
    fn <- sum(calls != HTO & donor_hto_list[seurat_object$genetic_donor] == HTO) #False negative rate
    f <- c(f, tp / (tp + 0.5 * (fp + fn)))
  }
#  f <- c(f, median(f)) #Add median F score
  f <- c(f, mean(f)) #Add mean F score

  names(f) <- c(donor_hto_list, "Average")
  
  return(f)
}
```

```{r}
Fscore_hashedDrops_b1 <- calculate_HTO_fscore(seu_b1, donor_hashtag_list_b1[1:4], "hashedDrops_calls")
Fscore_hashedDrops_b2 <- calculate_HTO_fscore(seu_b2, donor_hashtag_list_b2[1:4], "hashedDrops_calls")

Fscore_hashedDrops_default_b1 <- calculate_HTO_fscore(seu_b1, donor_hashtag_list_b1[1:4], "hashedDrops_default_calls")
Fscore_hashedDrops_default_b2 <- calculate_HTO_fscore(seu_b2, donor_hashtag_list_b2[1:4], "hashedDrops_default_calls")

Fscore_hashsolo_b1 <- calculate_HTO_fscore(seu_b1, donor_hashtag_list_b1[1:4], "hashsolo_calls")
Fscore_hashsolo_b2 <- calculate_HTO_fscore(seu_b2, donor_hashtag_list_b2[1:4], "hashsolo_calls")

Fscore_HTODemux_b1 <- calculate_HTO_fscore(seu_b1, donor_hashtag_list_b1[1:4], "HTODemux_calls")
Fscore_HTODemux_b2 <- calculate_HTO_fscore(seu_b2, donor_hashtag_list_b2[1:4], "HTODemux_calls")

Fscore_GMMDemux_b1 <- calculate_HTO_fscore(seu_b1, donor_hashtag_list_b1[1:4], "GMMDemux_calls")
Fscore_GMMDemux_b2 <- calculate_HTO_fscore(seu_b2, donor_hashtag_list_b2[1:4], "GMMDemux_calls")

Fscore_deMULTIplex_b1 <- calculate_HTO_fscore(seu_b1, donor_hashtag_list_b1[1:4], "deMULTIplex_calls")
Fscore_deMULTIplex_b2 <- calculate_HTO_fscore(seu_b2, donor_hashtag_list_b2[1:4], "deMULTIplex_calls")

Fscore_BFF_raw_b1 <- calculate_HTO_fscore(seu_b1, donor_hashtag_list_b1[1:4], "BFF_raw_calls")
Fscore_BFF_raw_b2 <- calculate_HTO_fscore(seu_b2, donor_hashtag_list_b2[1:4], "BFF_raw_calls")

Fscore_BFF_cluster_b1 <- calculate_HTO_fscore(seu_b1, donor_hashtag_list_b1[1:4], "BFF_cluster_calls")
Fscore_BFF_cluster_b2 <- calculate_HTO_fscore(seu_b2, donor_hashtag_list_b2[1:4], "BFF_cluster_calls")

Fscore_demuxmix_b1 <- calculate_HTO_fscore(seu_b1, donor_hashtag_list_b1[1:4], "demuxmix_calls")
Fscore_demuxmix_b2 <- calculate_HTO_fscore(seu_b2, donor_hashtag_list_b2[1:4], "demuxmix_calls")
```

```{r}
Fscore_matrix_b1 <- data.frame("Hashtag" = c(hashtag_list_b1[1:4], "Mean"),
                            "hashedDrops" = Fscore_hashedDrops_b1,
                            "hashedDrops_default" = Fscore_hashedDrops_default_b1,
                            "HashSolo" = Fscore_hashsolo_b1,
                            "HTODemux" = Fscore_HTODemux_b1,
                            "GMM_Demux" = Fscore_GMMDemux_b1,
                            "deMULTIplex" = Fscore_deMULTIplex_b1,
                            "BFF_raw" = Fscore_BFF_raw_b1,
                            "BFF_cluster" = Fscore_BFF_cluster_b1,
                            "demuxmix" = Fscore_demuxmix_b1)

Fscore_matrix_b1

#Removing average information
Fscore_matrix_b1 = Fscore_matrix_b1[1:4,]

```
```{r}
Fscore_matrix_b2 <- data.frame("Hashtag" = c(hashtag_list_b2[1:4], "Mean"),
                            "hashedDrops" = Fscore_hashedDrops_b2,
                            "hashedDrops_default" = Fscore_hashedDrops_default_b2,
                            "HashSolo" = Fscore_hashsolo_b2,
                            "HTODemux" = Fscore_HTODemux_b2,
                            "GMM_Demux" = Fscore_GMMDemux_b2,
                            "deMULTIplex" = Fscore_deMULTIplex_b2,
                            "BFF_raw" = Fscore_BFF_raw_b2,
                            "BFF_cluster" = Fscore_BFF_cluster_b2,
                            "demuxmix" = Fscore_demuxmix_b2)

Fscore_matrix_b2

#Removing average information
Fscore_matrix_b2 = Fscore_matrix_b2[1:4,]
```
```{r}
Fscore_matrix_b1 %>%
  pivot_longer(cols = c("hashedDrops", "hashedDrops_default", "HashSolo",
                        "HTODemux", "GMM_Demux", "deMULTIplex", 
                        "BFF_raw", "BFF_cluster", "demuxmix"),
               names_to = "method",
               values_to = "Fscore") -> Fscore_matrix_b1

Fscore_matrix_b2 %>%
  pivot_longer(cols = c("hashedDrops", "hashedDrops_default", "HashSolo",
                        "HTODemux", "GMM_Demux", "deMULTIplex", 
                        "BFF_raw", "BFF_cluster", "demuxmix"),
               names_to = "method",
               values_to = "Fscore") -> Fscore_matrix_b2
```



```{r}
Fscore_matrix_b1$Batch <- "Batch 1"
Fscore_matrix_b2$Batch <- "Batch 2"

Fscore_matrix <- bind_rows(Fscore_matrix_b1,
                           Fscore_matrix_b2)
```

```{r}
Fscore_matrix %>%
 group_by(Batch) %>%
  heatmap(.row = method, 
        .column = Hashtag, 
        .value = Fscore,
        column_title = "F-score - solid tissue data",
        cluster_rows = TRUE,
        row_names_gp = gpar(fontsize = 10),
        show_row_dend = FALSE,
        row_names_side = "left", 
        row_title = "",
        cluster_columns = FALSE,
        column_names_gp = gpar(fontsize = 10),
        palette_value = plasma(3)) -> p1

p1 <- wrap_heatmap(p1)
p1
```
```{r}
#  ggsave(here("paper_latex", "figures", "OT_Fscore.png"),
#       p1,
#       device = "png",
#       width = 8, height = 5,
#       units = "in",
#       dpi = 350
#      )
```

## Doublet assignments

Also worried about whether any of these methods are assigning genetic doublets as singlets. Going to look at the fraction of genetic doublets that get assigned to each of the possible HTO categories.

```{r}
b1_doublets <- seu_b1[, seu_b1$genetic_donor == "Doublet"]
b2_doublets <- seu_b2[, seu_b2$genetic_donor == "Doublet"]

b1_doublet_doublet <- NULL
b1_doublet_negative <- NULL
b1_doublet_singlet <- NULL

b2_doublet_doublet <- NULL
b2_doublet_negative <- NULL
b2_doublet_singlet <- NULL

for (method in method_calls) {
  b1_doublet_doublet <- c(b1_doublet_doublet, sum(seu_b1$genetic_donor == "Doublet" & seu_b1[[method]] == "Doublet") / sum(seu_b1$genetic_donor == "Doublet"))
  b1_doublet_negative <- c(b1_doublet_negative, sum(seu_b1$genetic_donor == "Doublet" & seu_b1[[method]] == "Negative") / sum(seu_b1$genetic_donor == "Doublet"))
  b1_doublet_singlet <- c(b1_doublet_singlet, sum(seu_b1$genetic_donor == "Doublet" & Reduce("|", lapply(hashtag_list_b1[1:4], function(x) seu_b1[[method]] == x))) / sum(seu_b1$genetic_donor == "Doublet"))
  
  b2_doublet_doublet <- c(b2_doublet_doublet, sum(seu_b2$genetic_donor == "Doublet" & seu_b2[[method]] == "Doublet") / sum(seu_b2$genetic_donor == "Doublet"))
  b2_doublet_negative <- c(b2_doublet_negative, sum(seu_b2$genetic_donor == "Doublet" & seu_b2[[method]] == "Negative") / sum(seu_b2$genetic_donor == "Doublet"))
  b2_doublet_singlet <- c(b2_doublet_singlet, sum(seu_b2$genetic_donor == "Doublet" & Reduce("|", lapply(hashtag_list_b2[1:4], function(x) seu_b2[[method]] == x))) / sum(seu_b2$genetic_donor == "Doublet"))
}

names(b1_doublet_doublet) <- method_calls
names(b1_doublet_negative) <- method_calls
names(b1_doublet_singlet) <- method_calls

names(b2_doublet_doublet) <- method_calls
names(b2_doublet_negative) <- method_calls
names(b2_doublet_singlet) <- method_calls
```

```{r}
b1_doublet_assignments <- data.frame("method" = c("hashedDrops", 
                                                  "hashedDrops (default)", 
                                                  "HashSolo", 
                                                  "HTODemux",
                                                  "GMM-Demux", 
                                                  "deMULTIplex", 
                                                  "BFF_raw",
                                                  "BFF_cluster", 
                                                  "demuxmix"),
                            "Doublet" = b1_doublet_doublet,
                            "Negative" = b1_doublet_negative,
                            "Singlet" = b1_doublet_singlet) %>%
  pivot_longer(cols = c("Doublet", "Negative", "Singlet"), 
               names_to = "assignment",
               values_to = "fraction")

b2_doublet_assignments <- data.frame("method" = c("hashedDrops", 
                                                  "hashedDrops (default)", 
                                                  "HashSolo", 
                                                  "HTODemux",
                                                  "GMM-Demux", 
                                                  "deMULTIplex", 
                                                  "BFF_raw",
                                                  "BFF_cluster", 
                                                  "demuxmix"),
                            "Doublet" = b2_doublet_doublet,
                            "Negative" = b2_doublet_negative,
                            "Singlet" = b2_doublet_singlet) %>%
  pivot_longer(cols = c("Doublet", "Negative", "Singlet"), 
               names_to = "assignment",
               values_to = "fraction")
```

```{r}
doublet_colours <- c("black", "gray60", "firebrick1")

p2 <- ggplot(b1_doublet_assignments %>%
               mutate(method = factor(method, levels = c("BFF_cluster",
                                                         "BFF_raw", 
                                                         "deMULTIplex",
                                                         "hashedDrops (default)", 
                                                         "demuxmix", 
                                                         "HTODemux", 
                                                         "GMM-Demux",
                                                         "HashSolo", 
                                                         "hashedDrops")))) +
  geom_bar(aes(x = method, y = fraction, fill = assignment),
           stat = "identity") +
  ggtitle("Batch 1 (1048 doublets)") +
  ylim(0, 1) +
  scale_fill_manual(values = doublet_colours) +
  theme(axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_text(size = 12)) + 
  coord_flip() + NoLegend()

p3 <- ggplot(b2_doublet_assignments %>%
               mutate(method = factor(method, levels = c("BFF_cluster",
                                                         "BFF_raw", 
                                                         "deMULTIplex",
                                                         "hashedDrops (default)", 
                                                         "demuxmix", 
                                                         "HTODemux", 
                                                         "GMM-Demux",
                                                         "HashSolo", 
                                                         "hashedDrops")))) +
  geom_bar(aes(x = method, y = fraction, fill = assignment),
           stat = "identity") +
  ggtitle("Batch 2 (805 doublets)") +
  ylim(0, 1) +
  scale_fill_manual(values = doublet_colours) +
  theme(axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()
        ) + coord_flip()

p2 | p3
```

```{r}
#  ggsave(here("paper_latex", "figures", "OT_doublet_assignments.png"),
#       p2 | p3,
#       device = "png",
#       width = 8, height = 5,
#       units = "in",
#       dpi = 350
#      )
```





